---
title: "STAT545_hmk2_WilliamHall"
author: "William Hall"
date: '2018-09-18'
output: html_document
---

```{r setup, include=FALSE, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = FALSE, include = FALSE}
#install.packages("gapminder")
#install.packages("tidyverse")
library(gapminder)
install.packages(ggpubr)
library(tidyverse)
library(ggplot2)
library(ggpubr)
```

## Smell Test the data

>Is gapminder a data.frame, a matrix, a vector, a list? What is gapminder's class?

```{r}
str(gapminder)
```

Using str(), we can see that gapminder has multiple classes 'tbl_df', 'tbl', and 'data.frame'. I think this means that it is a tibble/data frame.


>How many variables/columns does gapminder have? How many rows/observations does gapminder have?

Gapminder has `r nrow(gapminder)` rows, and `r ncol(gapminder)` columns.


>Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?

Yup. You could use anyone of the following functions to get information about the dataset.
  - names()
  - head()
  - summary()
  - str()
  - nrow()
  - ncol()

>What data type is each variable?

  - Country         is a `r class(gapminder$country)` variable.
  - Continent       is a `r class(gapminder$continent)` variable.
  - Year            is a `r class(gapminder$year)` variable.
  - Life Expectancy is a `r class(gapminder$lifeExp)` variable.
  - Population      is a `r class(gapminder$pop)` variable.
  - GDP             is a `r class(gapminder$gdpPercap)` variable.

## Life Expectancy Exploration

  - What are possible values (or range, whichever is appropriate) of each variable?
  - What values are typical? What’s the spread? What’s the distribution?
  
```{r}
summary(gapminder$lifeExp)
```

  - The minimum value of Life Expectancy is  `r min(gapminder$lifeExp)`.
  - The maximum value of Life Expectancy is  `r max(gapminder$lifeExp)`.
  - The median  value. of Life Expectancy is `r median(gapminder$lifeExp)`.
  - The mean value. of Life Expectancy is    `r mean(gapminder$lifeExp)`.
  
```{r}
ggplot(data=gapminder, aes(x=gapminder$lifeExp)) +
         geom_histogram(bins = 30, color = "black", fill = "gray") +
         geom_vline(aes(xintercept = mean(gapminder$lifeExp)), 
             linetype = "dashed", size = 0.6) +
         xlab("Life Expectancy") + 
         ylab("Count") + 
         ggtitle("Life Expectancy Histogram")
            
```

```{r}
ggplot(data=gapminder, aes(x=gapminder$lifeExp)) +
  geom_density() +
  geom_vline(aes(xintercept = mean(gapminder$lifeExp)), 
             linetype = "dashed", size = 0.6)
  xlab("Life Expectancy") + 
  ylab("Density") + 
  ggtitle("Life Expectancy Density Plot")

```

## Continent Exploration

The Contient variable in the Gapminder dataset is a `r class(gapminder$continent)` variable. It has the following levels `r levels(gapminder$continent)`.

```{r}
levels(gapminder$continent)
```

The following describes the frequency of each level.

```{r}
summary(gapminder$continent)
```

```{r}
ggplot(gapminder, aes(continent)) +
  geom_bar(fill = "#0073C2FF") +
  ggtitle("Frequency of Continents in Gapminder Dataset")
```

##Explore Plotting

> Scatterplot of two quantitative variables

```{r}
ggplot(gapminder, aes(x=gdpPercap, y=lifeExp)) +
  geom_point() +
  xlab("GDP Per Cap") + 
  ylab("Life Expectancy") + 
  ggtitle("Life Expectancy and GDP")

```

> Plot of one quantitative variable

```{r}
ggplot(data=gapminder, aes(x=gapminder$lifeExp)) +
         geom_histogram(bins = 30, color = "black", fill = "gray") +
         geom_vline(aes(xintercept = mean(gapminder$lifeExp)), 
             linetype = "dashed", size = 0.6) +
         xlab("Life Expectancy") + 
         ylab("Count") + 
         ggtitle("Life Expectancy Histogram")
```

> Plot of one quantitative and one categorical variable

```{r}
ggplot(gapminder, aes(continent, lifeExp)) +
  geom_boxplot() +
  xlab("Continent") + 
  ylab("Life Expectancy") + 
  ggtitle("Life Expectancy for Continents")
```


## Show me your PIPES

> So what if we wanted to compare the mean GDP for countries in Africa with life expectancy.

```{r}

gapminder %>% 
    filter(continent == "Africa") %>% 
    group_by(country) %>% 
    summarize(Mean.Life = mean(lifeExp), GDP = mean(gdpPercap))

gapminder %>% 
  filter(continent == "Africa") %>% 
  ggplot(aes(lifeExp, gdpPercap)) +
  geom_point(aes(col = country))

          
```


##Reflection

This assignment was fun. One thing I struggle with is the %>% pipe vs. the + for ggpplot. I figured it out eventually, but I definitely lost some hair trying to figure out what was wrong with my code. I can see how the tidyverse is super powerful, and why everyone loves it so much :)

Can't wait to dig into some more of the advanced features!